### 2017-09-06 conclusions.tex
moving staff

\section{Conclusion and Future Work}

Our proposed technique for relaxed instrumentation can also be applied for similar profiling scenarios, such as profiling empirical computational complexity~\cite{goldsmith07,zaparanuks12,coppa14}.