### 2017-09-06 premable.tex

\usepackage{booktabs} % For formal tables

\usepackage[utf8]{inputenc}
\usepackage{fullpage}

\usepackage{graphics}

\usepackage{graphicx}
\usepackage{subcaption}

\usepackage{multirow}

%\usepackage{setspace}
\usepackage{ragged2e}

	\usepackage{soul}

%\usepackage{flushend}

%\usepackage{indentfirst}

\usepackage{amsthm}
\usepackage{amsmath,amssymb}
%\usepackage{mathrsfs}
\usepackage{mathtools}

\theoremstyle{definition}
\newtheorem{prop}{Proposition}[section]

%% Listing related stuff

\usepackage{listings}
\usepackage{courier}            %Required for pretty listings (more dense)

%% EMACS colors for listing:

\usepackage{color}
\definecolor{sh_comment}{rgb}{0.65, 0.00, 0.00 } % red-ish
\definecolor{sh_keyword}{rgb}{0.37, 0.69, 0.69}  % blue-green
\definecolor{sh_string}{rgb}{0.08, 0.69, 0.08} % green

%% This manipulates the listing numbering. It decreases it by one at the point used.
\newcommand*\lstDNumber{\addtocounter{lstnumber}{-1}}

%% Change from Listing->Algorithm
%% \renewcommand\lstlistingname{Algorithm} % Fix listing caption name: Listing -> Algorithm
%% \renewcommand\lstlistlistingname{Algorithms}
%% \def\lstlistingautorefname{Alg.}

%% \renewcommand{\thelstlisting}{\thesection-\arabic{lstlisting}}
%\renewcommand{\thelstlisting}{\arabic{lstlisting}} % Fix listing numbering Algorithm 1.1 -> Algorithm 1.

%% \renewcommand{\labelenumi}{\roman{enumi}.}

\lstset{
float=[*],
language=C,                % choose the language of the code
%% basicstyle=\linespread{0.9}
%% basicstyle=\scriptsize\ttfamily,
 basicstyle={\linespread{0.97} \scriptsize\ttfamily},
 stringstyle=\color{sh_string},
 keywordstyle = \color{sh_keyword}\bfseries,
 commentstyle=\color{sh_comment}\itshape,
numbers=left,                   % where to put the line-numbers
%% numberstyle=\scriptsize,        % the size of the fonts that are used for the line-numbers
numberstyle=\scriptsize,        % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
xleftmargin=2em,                % Left margin
frame=lines,                   % adds a frame around the code. none|leftline|topline|bottomline|lines|single|shadowbox
framexleftmargin=1.5em,         % Margin from frame to line numbers
framexbottommargin=0em,         % Distance from last text to frame
morekeywords={in,true,false,and,or,set,is,not},
%% prebreak=\mbox{\tiny$\searrow$},
prebreak=\space,                % Line break: insert this character at the end of the top line
%% postbreak=\mbox{{\color{blue}\scriptsize$\hookrightarrow$}}, % Line break: blue arrow at the beginning of the bottom line
postbreak=\mbox{{\color{blue}\scriptsize$\hookrightarrow$}}, % Line break: blue arrow at the beginning of the bottom line
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to top
escapeinside={\@}{\@}             % Escape sequence \@ESCAPED STUFF\@
}

\usepackage{nasm/lang}
\usepackage{nasm/style}
\usepackage{llvm/lang}

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{10.475/123_4}

% ISBN
%\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[CGO'18]{International Symposium on Code Generation and Optimization}{2018}{}
%\acmYear{1997}
%\copyrightyear{2016}
%
%\acmPrice{15.00}

\newcommand{\etal}{et~al.}

\newcommand{\itercomp}{{iterative compilation}}
\newcommand{\Itercomp}{{Iterative compilation}}
\newcommand{\IterComp}{{Iterative Compilation}}
\newcommand{\flagstype}{\usefont{T1}{cmr}{m}{n}}