### 2017-09-14 conclusions.tex
first attack on conclusions

\section{Conclusions}
This paper has presented a novel work efficiency metric to enable one to compare the performance of any two different compiled versions of
a program without executing a given input more than once. This ability allows us to implement, for the first time, a true online iterative
compilation  to search for the best compilation options  based on actual user inputs seen in the deployment environment.

We develop a compiler-based, low-overhead instrumentation mechanism to gather the require data to evaluate the work metric. Our mechanism
does this without any human involvement. We show that by carefully selecting where to place the probes in the source code, the
instrumentation overhead can be negligible. We give two probe optimization strategies, allowing developers to balance the instrumentation
overhead and the performance error.

We  evaluate our approach using the KDataSets benchmark suit. Experimental results show that our approach delivers \red{80\%} of the
speedup achieved by an offline  oracle, but without executing any input more than once. We show that our probe optimization strategy is
highly effective, which brings the instrumentation overhead down to only \red{4\%} on average, making our approach practical for regular
use.

%Our proposed technique for relaxed instrumentation can also be applied for similar profiling scenarios, such as profiling empirical
%computational complexity~\cite{goldsmith07,zaparanuks12,coppa14}.