### 2017-09-14 cgo18mscr.tex
Splitting

\PassOptionsToPackage{table}{xcolor}

\documentclass[sigplan,10pt]{acmart}
\input{premable}

\begin{document}
\title{Online {\IterComp}}

\author{  }

%\author{Rodrigo C. O. Rocha}
%\affiliation{%
%  \institution{University of Edinburgh, UK}
%}
%\email{r.rocha@ed.ac.uk}

%\author{Pavlos Petoumenos}
%\affiliation{%
%  \institution{University of Edinburgh, UK}
%}
%\email{ppetoume@inf.ed.ac.uk}

%\author{Lu\'is F. W. G\'oes}
%\affiliation{%
%  \institution{PUC Minas, Brazil}
%}
%\email{lfwgoes@pucminas.br}

%\author{Murray Cole}
%\affiliation{
%  \institution{University of Edinburgh, UK}
%}
%\email{mic@inf.ed.ac.uk}

%\author{Zheng Wang}
%\affiliation{%
%  \institution{Lancaster University, UK}
%}
%\email{z.wang@lancaster.ac.uk}

%\author{Hugh Leather}
%\affiliation{%
%  \institution{University of Edinburgh, UK}
%}
%\email{hleather@inf.ed.ac.uk}

%\renewcommand{\shortauthors}{R. Rocha et al.}

\input{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10011007.10011006.10011041</concept_id>
%<concept_desc>Software and its engineering~Compilers</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%<concept>
%<concept_id>10011007.10011006.10011041.10011043</concept_id>
%<concept_desc>Software and its engineering~Retargetable compilers</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Software and its engineering~Compilers}
%\ccsdesc[500]{Software and its engineering~Retargetable compilers}
\keywords{{\IterComp}, Profiling} %, Relaxed Instrumentation}

\maketitle
\vspace{-1em}

\input{intro}
\input{motivation}
\input{work}
\input{infra}
\input{prof}
\input{setup}
\input{results}
\input{related}
\input{conclusions}

%
%\section*{Acknowledgments}
%
%This work was supported by the UK Engineering
%and Physical Sciences Research Council (EPSRC) under grants
%EP/L01503X/1 for the University of Edinburgh, School
%of Informatics, Centre for Doctoral Training in Pervasive
%Parallelism, % (\url{http://pervasiveparallelism.inf.ed.ac.uk/}),
%and also by the Institute for Computing Systems Architecture (ICSA)
%in the School of Informatics at the University of Edinburgh.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}