### 2017-09-16 related.tex
done with related work

\section{Related Work}\label{sec:relatedwork}

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03,hoste08}. These approaches compare the performance of candidate compiler options by first
compile the program with different options and then profiling the different compiled versions on the same program input. All these
approaches implicitly assume that the best-performing compiler settings found from a small set of inputs will work well on unseen datasets.
Chen~\etal~\cite{chen10,chen12a} show that the best optimization options can differ across inputs. Their results suggest that, in most
cases it is possible to find a single set of optimizations that produces good results across all inputs; but achieving this requires
developers provide a few tens of representative inputs to be used by the iterative compilation search process. We argue that providing
representative inputs is difficult in practice, as this require developers to foresee all possible user inputs. Even this is possible,
current approaches all require to evaluate each candidate compiler option on every single input. Doing so incurs significant overhead,
which thus prevents iterative compilation to be adopted at a larger scale.

%=======
%Multiple adaptive compilation techniques have been proposed for finding a near optimal combination of optimization
%settings~\cite{agakov06,hoste08,kulkarni04,stephenson03} by probabilistically searching the space of possible combinations. But with
%all of them there is uncertainty about whether the quality of their results depends on the data set used. Chen~\etal~\cite{chen10,chen12a}
%examined this issue and showed that indeed the best optimization options differ across inputs. Still, in most cases it is possible to find
%an optimization sequence that produces good results across all inputs, by evaluating the optimizations on a representative set inputs.
%While a positive result, this is still impractical. Producing a set of realistic inputs might be difficult. Even if not, each
%optimization sequence has to be evaluated on every input, increasing the time \itercomp takes by at least an order of
%magnitude.
%>>>>>>> origin/master

Recent studies~\cite{chen12b,fang15} have used real data to perform {\itercomp} on distributed data center applications. Each worker
receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from each round is used
for subsequent executions of the same code. By testing new settings and re-evaluating old ones on new datasets this approach makes it more
likely to select an optimization setting that works well across inputs. Like in~\cite{chen10} every setting is tested on the same set of
inputs, wasting time and energy. This approach only works well with MapReduce-like workloads, since it relies on the MapReduce framework
for repeating the same computation multiple times without side-effects or corrupting the input.

Fursin~\etal~\cite{fursin07} attempted to find a general approach for online \itercomp. They proposed to use instructions per cycle (IPC)
as a metric for comparing different runs of the same program with different optimization settings and different inputs. Their hypothesis is
that application binaries displaying higher IPC are more efficient than lower IPC binaries. However, many optimization techniques affect
IPC in the opposite way, making the program more efficient while lowering IPC. We have also shown in Section~\ref{sec:motivation}, IPC is
not strongly correlate to speedup. Therefore, it is not a good metric for \itercomp. Furthermore, IPC cannot be used to estimate
performance for multi-threaded applications~\cite{alameldeen06,eyerman08} due to the side-effect of thread contention. Aalameldeen and
Wood~\cite{alameldeen06} also suggest that relying on IPC can be harmful, and work metrics like time per work unit or work units per time
should be used. However, prior works all rely on developers to provide such metrics. This paper presents the first automatic approach to
derive such work metrics from arbitrary code without developer intervention.

%\subsection{Work and Input Size Metrics}

%Previous work have proposed profiling-based mechanism to estimate input sizes~\cite{zaparanuks12,coppa14}.
%Coppa~\etal~\cite{coppa14} in particular propose the concept of \textit{read memory size} for automatically estimating the size of the input passed to a routine, where \textit{read memory size} represents the number of distinct memory cells first accessed by a read operation.
%In other words, the \textit{read memory size} metric measures the size of the useful portion of the input's memory footprint.
%However, because we are interested in the amount of computational work performed in respect of a given input, the memory footprint of the input may not always have a direct correspondence to  the amount of computational work.

%Goldsmith~\etal~\cite{goldsmith07} use \textit{block frequency} as the measure for performance for empirically describing the asymptotic behavior of programs, which is known as empirical computational complexity.
%Block frequency is a relative metric that represents the number of times a basic block executes~\cite{ball94,ball96}.
%They argue in favor of block frequency due to its portability, repeatability and exactness, since it does not suffer from timer resolution problems or non-deterministic noises.
%Block frequency also has the advantage of being efficiently profiled by means of automatic code instrumentation~\cite{knuth73,ball94}.

%However, in the context of comparing different optimizations, although block frequency would be able to capture aspects of optimizations that simplify the control-flow graph (CFG), measuring work at the basic block resolution would not capture effects of optimizations at the instruction level.
%Because of that, we extend the idea of using basic block frequency to measure computational work by also considering the computational cost of each basic block.
%The computational cost of a basic block is given by weighing the instructions that it contains.

\subsection{Optimal Instrumentation}

The seminal work of Donald Knuth~\cite{knuth73} introduced an optimal algorithm for profiling block frequency, inserting much fewer probes
than the na\"ive instrumentation approaches that were used in practice~\cite{knuth71}. Forman~\cite{forman81} proposed an approach to
further reduce the instrumentation overhead by placing probes in basic blocks that are less likely to be executed based on static
heuristics. Ball and Larus~\cite{ball94} offered a detailed discussion to compare these two approaches for optimally profiling block
frequency, namely, placing probes on the vertices or the edges of the control-flow graph (CFG). They show that the edge-based approach
produces better placement of probes. Our work built upon these prior works. We use the edge-based approach to place our probes before
applying relaxation.