### 2017-11-02 motivation.tex
Alternative motivation section. Old one saved in another file

% vim: set tabstop=4 shiftwidth=4 expandtab : 
\section{Motivation \label{sec:motivation}}

    % Representative inputs are important. Example.
    % Artificial representative input sets are hard to create
    % Representative input sets created out of real data are also hard to create
    % Just using real data online is not enough. Efficiency metrics are needed
    % We do not have a proper efficiency metric. Example.
    % The paper shows how to construct one.

    In this section we first motivate online iterative compilation by showing the deleterious effects of using non-representative input
    sets and that it is hard to create representative ones. Then we go on to examine what stops us from just using real inputs online and
    what we need to overcome the problem.

    \paragraph{Iterative compilation needs representative inputs.} \FIXME{Fig.XXX} shows what happens when we apply iterative compilation
    on \FIXME{susan\_c} using the wrong inputs. We selected \FIXME{50} of the 1,000 inputs of \textit{KDataSets}~\cite{chen10,chen12a} for
    \FIXME{susan\_c} and we used them to evaluate different optimizations. The red bars show the average speedup over \texttt{-O3} achieved
    by the optimization sequence selected by iterative compilation. On the left is the average over the \FIXME{50} inputs, on the right the
    average over all 1,000 inputs. The selected sequence works great for the inputs we used, but it fails miserably for the rest, slowing
    down the benchmark by \FIXME{5\%} on average.

    The blue bars, on the other hand, show the ideal case where good optimizations are selected based on their average effect on all
    inputs, not just a subset. Not only is the average speedup over all 1,000 inputs considerable, \FIXME{13\%} over \texttt{-O3}, but it
    also works well for the \FIXME{50} input subset. The speedup for the subset is only \FIXME{1\%} less of what we got when we optimized
    specifically for that subset.

    Making iterative compilation work in this example is not a matter of handling input sensitivity, we were able to find an optimization
    sequence that works well both on average and under most inputs. It is a matter of finding a set of inputs that represents the full
    range of inputs the application is likely to process. If we do, we can improve average performance by \FIXME{13\%}. If not, we might
    hurt average performance instead.

    \paragraph{Representative input sets are hard to create.} For developers to create artificial input sets, they first need to know how
    their application is actually being used. This is not always easy or even possible. Then, they need to create data that will trigger a
    similar behavior and this needs a deep knowledge of the application and considerable effort. This will have to be a continuous work in
    progress, to handle changes in the way the application is used.

    Collecting real input data online and processing them again offline to evaluate optimizations is not a solution either. It requires
    manual modifications to the application to save the data, more modifications to avoid side effects while reusing data, and even more
    modifications to ensure that the application does the same work every time the same data are reused, despite changes in the rest of
    the system state. This is significant engineering work. Even if this was not the case, many applications are just not able to save
    data for privacy or storage overhead reasons.

    \paragraph{Using real inputs online is not enough.} We could overcome these problems by just evaluating optimization sequences online,
    while the application is being used. This gives us real data but creates a new problem. Iterative compilation finds good optimizations
    by measuring the efficiency of the resulting binary in terms of speedup over another baseline binary. This requires processing the same
    input in the exact same way twice, one with the optimized and one with the baseline binary, measuring the runtimes, and calculating the
    speedup. When done online, we cannot process the same input twice, each time we process whatever the input happens to be. We need
    another efficiency metric that only requires a single use of an optimized binary.

    \paragraph{Existing efficiency metrics are not up to the task.} A application-specific ones~\cite{alameldeen06,coppa14} may work well,
    but the developer has to choose the appropriate metric and then manually modify the application to calculate it. Existing
    application-agnostic ones do not work. We run each of the \FIXME{32} \textit{KDataSets} benchmarks 1,000 times, each time with a
    different input \textit{and} a different optimized binary. We selected three performance metrics that can be easily measured in a
    single run and can potentially be used as efficiency metrics: runtime, inverse runtime, and
    \textit{Instructions Per Cycle}~(IPC)~\citep{fursin07}.

    \FIXME{Fig.XXX} shows how well these three candidates for an efficiency metric correlate with the metric we would actually like to have,
    speedup over \texttt{-O0}. For all benchmarks, the correlation coefficient~\cite{Bishop:2006:PRM:1162264} is below \FIXME{0.20} for
    runtime and below \FIXME{0.35} for inverse runtime. Both have obvious reasons for being inappropriate efficiency metrics. Doubling the
    amount of data to be processed will cause the runtime to rise and the inverse runtime to fall, even if we use the same binary.
    Efficiency has nothing to do with it. IPC also correlates poorly with performance speedup, the correlation coefficient being just
    \FIXME{0.15}. While it is a common efficiency metric in computer architecture, it works there because it is used with invariant binaries.
    In contrast, any compiler optimization that trades high latency instructions for multiple low latency instructions can reduce IPC while
    improving efficiency.

    No existing metric can give us an application-agnostic measure of efficiency with a single execution. The rest of the paper will show
    how to create one and how to use it for iterative compilation.