### 2017-11-09 cgo18mscr.tex
Switched template to PLDI 2018

\PassOptionsToPackage{table}{xcolor}
\PassOptionsToPackage{draft}{hyperref}
\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\input{preamble}

\begin{document}
\title{Online {\IterComp}}

\author{Rodrigo C. O. Rocha}
\affiliation{
  \institution{University of Edinburgh, UK}
}
\email{r.rocha@ed.ac.uk}

\author{Pavlos Petoumenos}
\affiliation{
  \institution{University of Edinburgh, UK}
}
\email{ppetoume@inf.ed.ac.uk}

\author{Lu\'is F. W. G\'oes}
\affiliation{
  \institution{PUC Minas, Brazil}
}
\email{lfwgoes@pucminas.br}

\author{Murray Cole}
\affiliation{
  \institution{University of Edinburgh, UK}
}
\email{mic@inf.ed.ac.uk}

\author{Zheng Wang}
\affiliation{
  \institution{Lancaster University, UK}
}
\email{z.wang@lancaster.ac.uk}

\author{Hugh Leather}
\affiliation{
  \institution{University of Edinburgh, UK}
}
\email{hleather@inf.ed.ac.uk}

\renewcommand{\shortauthors}{R. Rocha et al.}

\input{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011041</concept_id>
<concept_desc>Software and its engineering~Compilers</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011041.10011043</concept_id>
<concept_desc>Software and its engineering~Retargetable compilers</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Compilers}
\ccsdesc[500]{Software and its engineering~Retargetable compilers}
\keywords{{\IterComp}, Profiling} %, Relaxed Instrumentation}

\maketitle
\vspace{-1em}

\input{intro}
\input{motivation}
\input{work}
\input{infra}
%\input{prof}
\input{setup}
\input{results}
\input{related}
\input{conclusions}

%
%\section*{Acknowledgments}
%
%This work was supported by the UK Engineering
%and Physical Sciences Research Council (EPSRC) under grants
%EP/L01503X/1 for the University of Edinburgh, School
%of Informatics, Centre for Doctoral Training in Pervasive
%Parallelism, % (\url{http://pervasiveparallelism.inf.ed.ac.uk/}),
%and also by the Institute for Computing Systems Architecture (ICSA)
%in the School of Informatics at the University of Edinburgh.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}